Minimal trap spaces of Logical models are maximal siphons of their Petri net encoding

### 2022-05-06 cmsb22.tex
Answers and cleanup… will work on the abstract now

% \title{Minimal trap spaces of Logical models are maximal siphons of their Petri net encoding}
% \titlerunning{Minimal trap spaces as maximal siphons}
\title{Scaling-up attractor computation for Boolean models}

@@

We will briefly recall here some preliminaries on Boolean models related to trap spaces and Petri nets.
To stay in line with the usual considerations we will always assume purely Boolean logical models, and an asynchronous update.
In the case of multi-level models, an encoding into a Boolean model is always possible.
As far as different update strategies are concerned, we think that even if they might require small technical adaptations, they would not bring much change to the main results of this article.

@@

Note that transforming a Boolean model defined by its functions into its Petri net encoding roughly relies on obtaining the Disjunctive Normal Forms (DNF) of each Boolean function.
Though this might be quite computationally intensive it is important to remark first that contrary to the prime-implicants case, there is no need to find \emph{minimal} DNFs.
Second, in practice this transformation seems to behave quite well on even quite big models as will be shown in the Section~\ref{sec:eval} on evaluation.
Currently, relying on BioLQM implies going through a BDD encoding of the formulae, which might not be optimal.
Hence we are still considering a direct encoding that would skip the BDD construction.

@@

\sylvain{What do you mean by \emph{advanced}? It is less since there might be an exponential number of prime implicants?}

@@

Though we benchmarked this new approach against the state-of-the-art tool PyBoolNet, there are many more evaluations that we plan to do in the future.
First, the BioLQM platform that we are using is providing another implicant-based method using BDDs in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} and though we expect it to behave mostly like PyBoolNet, that remains to be checked.
Note that this also raises the question of replacing altogether the BioLQM preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.

>>>>

\title{Minimal trap spaces of Logical models are maximal siphons of their Petri net encoding}
\titlerunning{Minimal trap spaces as maximal siphons}
% \title{Scaling-up attractor computation for Boolean models}
% \vangiang{I like the first title :). Since we only conduct experiments on real-world Boolean models of not too large size, it seems to make less sense to use "scaling-up" (this should be for the journal version).}

@@

We will briefly recall here some preliminaries on Boolean models related to trap spaces and Petri nets.
In the case of multi-level Logical models, an encoding into a Boolean model is always possible.

@@

Note that transforming a Boolean model defined by its functions into its Petri net encoding roughly relies on obtaining conditions for the activation/inactivation of the states. In~\cite{chaouiya2004qualitative} this took the form of the whole truth table of the Boolean functions, but as shown in Appendix 1 of~\cite{chatain2014characterization} computing Disjunctive Normal Forms (DNF) of each Boolean function is enough.
Though this might appear quite computationally intensive it is important to remark first that contrary to the prime-implicants case, there is no need to find \emph{minimal} DNFs.
Second, in practice this transformation, here using BDDs in BioLQM, seems to behave quite well on even quite big models as will be shown in the Section~\ref{sec:eval} on evaluation.

@@

\sylvain{What do you mean by \emph{advanced}? It is less since there might be an exponential number of prime-implicants?}

@@

Though we benchmarked this new approach against the state-of-the-art tool PyBoolNet, there are many more evaluations that we plan to do in the future.
First, the BioLQM platform that we are using is providing another implicant-based method using BDDs in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} and though we expect it to behave mostly like PyBoolNet, that remains to be checked.
Note that this also raises the question of replacing altogether the BioLQM preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
The tradeoff between a small Petri net, and hence small ASP, and the time it takes to compute it has to be evaluated in depth.
